Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation\\ (Supplementary Materials)

### 2011-04-06 supplementary_material.tex
Typesets properly without appendix, spills with appendix.

\begin{IEEEkeywords}
Face Recognition, Face Alignment, Illumination Variation, Occlusion and Corruption,
Sparse Representation,  Error Correction, Validation and Outlier Rejection.
\end{IEEEkeywords}}

@@

%%%%%%%%% BODY TEXT
\section{Introduction}
%Automatic face recognition remains one of the most active areas in computer vision.

\PARstart{F}{ace} recognition applications to date have fallen
into roughly two categories.  Face recognition has recently
seen a lot of success in a family of less-demanding
applications such as online image search and family photo album
organization (e.g.\ Google Picassa, Microsoft Photo Gallery,
and Apple iPhoto). At the other end of the tractability
spectrum there are the terrorist watchlist and mass
surveillance applications that have for the most part dominated
the field of face recognition research.  However, there are
many face recognition applications that fall roughly between
these extremes, where very high recognition performance is
desired, but the users in the gallery are still allies of the
system rather than adversaries.  These applications include
access control for secure facilities (e.g., prisons, office
buildings), computer systems, automobiles, or automatic teller
machines, where controlled gallery images can be obtained in
advance.  These applications are very interesting due to their
potential sociological impact.  Since the gallery subjects are
allies, rather than opponents, of the recognition system, this
creates the possibility of carefully controlling the
acquisition of the training data. While the same can be said
for other biometrics such as fingerprints and iris recognition,
face recognition has the potential of working with test data
that is much less controlled, allowing the access control
system to be made less intrusive to the users of the system. To
some extent, the goal of this paper is to show how a reliable
face recognition system can be built for this restricted, but
still important, scenario.

Very few recognition systems specifically target applications where many well-
controlled training images are available.  Of these, the classical holistic
subspace-based face recognition methods \cite{Turk1991-CVPR,Belhumeur1997-PAMI}
are well known for their speed and simplicity, as well as for their natural
extension to linear illumination models.  However, their performance has been
shown to be extremely brittle not only to alignment variation, but to even
minor occlusions caused by, say, a wisp of hair, a blinked eye, or mouth that
is slightly open. One promising recent direction, set forth in
\cite{Wright2009-PAMI}, casts the recognition problem as one of finding a
sparse representation of the test image in terms of the training set as a
whole, up to some sparse error due to occlusion. A \emph{sparse
representation-based classification} (SRC) method is then proposed for
recognition. The main idea is that the sparse nonzero coefficients should
concentrate on the training samples with the same class label as the test
sample. SRC has demonstrated striking recognition performance despite severe
occlusion or corruption by solving a simple convex program.
%For this reason, the final recognition stage of our proposed system will
%consist of SRC, and the alignment stage that precedes it will be based on
%similar techniques.

\begin{figure}
\centering \begin{tabular}{cc}
\includegraphics[height=1in]{figures_cvpr/promo/case1/detector.png}&
\hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case1/sci_with_axis_face_case1.png}
\\ \includegraphics[height=1in]{figures_cvpr/promo/alignment_and_detector.png}&
\hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case2/sci_with_axis_face_case2.png}
\\ \includegraphics[height=1in]{figures_cvpr/promo/case3/alignment.png} &
\hspace{3mm}
\includegraphics[height=1in]{figures_cvpr/promo/case3/sci_with_axis_face_case3.png}
\end{tabular} \caption{{\bf Effects of registration and illumination on
Recognition}. In this example we identify the girl among 20 subjects, by
computing the sparse representation of her input face with respect to the
entire training set. The absolute sum of the coefficients associated with each
subject is plotted on the right. We also show the faces reconstructed with each
subject's training images weighted by the associated sparse coefficients. The
red line (cross) corresponds to her true identity, subject 12. {\bf Top:} The
input face is from Viola and Jones' face detector (the black box) and all 38
illuminations specified in Section \ref{sec:illumination} are used in the
training.  {\bf Middle:} The input face is well-aligned (the white box) with
the training by our algorithm specified in Section \ref{sec:registration} but
only 24 frontal illuminations are used in the training for recognition (see
Section \ref{sec:illumination}). {\bf Bottom:} The input face is well aligned and
a sufficient set (all 38) of
illuminations are used in the training. Both are necessary for correct recognition
using SRC.}\label{fig:promo}
\vspace{-6.9mm}
\end{figure}

Unfortunately, while the work \cite{Wright2009-PAMI} achieves impressive results, it does not deal with misalignment
between the test and training images, and it requires a rich set of illuminations
in the gallery images for good performance.
We illustrate the compounded effect of both alignment and illumination
with an example in Figure \ref{fig:promo}.
The task is to identify the girl among 20 subjects. If the test
face image, say obtained from an off-the-shelf face detector,
has even a small amount of registration error against the
training images (caused by mild pose, scale, or misalignment),
the sparse representation obtained using the method of \cite{Wright2009-PAMI} is no longer informative,
even if sufficient illuminations are present in the training, as shown in Figure
\ref{fig:promo}(top). Additionally, in order to span
the illuminations of a typical indoor (or outdoor) environment,
illuminations from behind the subject are needed in the
training set. Otherwise, even for perfectly aligned test images,
the sparse representation obtained using \cite{Wright2009-PAMI} will not necessarily be sparse or
informative, as shown by the example in Figure
\ref{fig:promo}(middle). Clearly, both good alignment, as well
as sufficient training images are needed to ensure success of the sparsity-based
recognition method proposed by \cite{Wright2009-PAMI}.
In this paper, we examine how to handle alignment and illumination simultaneously
in the sparse representation framework, bringing the method proposed in \cite{Wright2009-PAMI}
closer to practical use.

\vspace{-0mm}
\subsection{Related Work}
We briefly review existing techniques for recognition, image
registration, and handling of illumination variation.  Our
system is based purely on 2D techniques.  This fact immediately
distinguishes our approach from systems that either require a
3D data as an input, or attempt to estimate a 3D model from 2D
input \cite{amberg2007reconstructing,Blanz2003-PAMI}.  While
these techniques can achieve better robustness to pose
variation given a sufficiently accurate 3D model, for access
control applications where only moderate pose variation is
present, the proposed method will be more than sufficient.
Note that 2D images of faces under varying illuminations
already contain 3D shape-related information,\footnote{In
principle, one can recover the 3D shape of the face from
multiple illuminations using photometric stereo.} and this
information can be leveraged by 2D algorithms for alignment and
recognition even if shape is not reconstructed explicitly.

% Alignment
In holistic recognition algorithms, correspondence between points in the test
image and in the training must be achieved.  A long line of research exists on
using Active Appearance Models \cite{Cootes2001-PAMI}, and the closely related
Active Shape Models \cite{cootes1992active} to register images against a
relatively high-dimensional model of plausible face appearances, often
leveraging face-specific contours.  While these model-based techniques have
advantages in dealing with variations in expression and pose, they may add
unnecessary complexity to applications where subjects normally present a neutral face
or only have moderate expression. We prefer to focus on deformations with far fewer
degrees of freedom, i.e. similarity transformations, and to use the training images themselves
as the appearance model. Iterative registration in this spirit dates at least back to the
Lucas-Kanade algorithm \cite{lucas1981iterative}.

Whereas much of the early work on image registration is aimed at the
problem of registering nearly identical images, say by minimizing a sum of
squared distances or maximizing normalized correlation, here we must confront several physical factors
simultaneously: misalignment, illumination variations, and corrupted pixels.
As we discuss further below, illumination variation can be dealt by
expressing the test image as a linear combination of an appropriate set of
training images. Similar representations have been exploited in
illumination-robust tracking (e.g., \cite{Belhumeur1999-PAA,Murase1995-IJCV}).
For robustness to gross errors, the $\ell^1$-norm of the residual is a more
appropriate objective function than the classical $\ell^2$-norm. Its use here
is loosely motivated by theoretical results due to Cand\`{e}s and Tao
\cite{CandesE2005-IT} (see also \cite{Wright2008-IT}). These two observations
lead us to pose the registration problem as the search for a set of
transformations and illumination coefficients that minimize the $\ell^1$-norm
of the representation error. We solve this problem using a generalized
Gauss-Newton method which solves a sequence of affine-constrained $\ell^1$-norm
minimization problems \cite{Osborne1990-JAMSSB,Jittorntrum1980-NM}. Each of
these problems can also be solved efficiently using recently developed
first-order techniques for $\ell^1$-minimization, which are reviewed in
\cite{YangA2010-pp}.

% Illumination
Researchers have tried various techniques to deal with illumination variation.
In almost all recognition algorithms where only a single gallery image is
available per individual, illumination effects are regarded as a nuisance that
must be removed before the algorithm can continue.  This is typically done by
making statistical assumptions about how illumination affects the image, and
using those assumptions to extract a new representation that is claimed to be
illumination invariant.  Recent examples include \cite{chen2006total} and
\cite{zhou2007appearance}.  However, despite these efforts, truly
illumination-invariant features are difficult to obtain from a single input
image.  We argue that if one has the luxury of designing the acquisition system
and the application demands a high recognition rate,
it is then unwise to limit the gallery to a
single image per person.  We therefore take the strategy of sampling many
gallery images of each individual under varying illuminations.  These images
are used as the basis for either a convex cone model
\cite{Georghiades2001-PAMI,belhumeur1998set}, or a subspace model
\cite{Basri2003-PAMI}.  Images are captured using a simple-to-construct
projector based light stage.  While similar systems have been used for
other applications, to our knowledge, we
are the first to use projectors to indirectly illuminate a subject's face for
the purpose of face recognition.
\subsection{Contributions} In this paper, we show how
registration and illumination
can be simultaneously addressed within a robust sparse representation
framework. We show that face registration, a challenging
nonlinear problem, can be solved by a series of linear programs
that iteratively minimize the sparsity of the registration
error. This leads to an efficient and effective alignment
algorithm for face images that works for a large range of
variation in translation, rotation, and scale, even when
the face is only partially visible due to eyeglasses, closed
eyes and open mouth, sensor saturation, etc.  We also propose a
sufficient set of training illuminations
that is capable of linearly representing typical indoor and outdoor
lighting, along with a practical hardware system for capturing
them.

% The two main contributions of this paper are thus to propose a simple and practical face recognition algorithm that achieves a high degree of robustness and stability to all these variations given sufficient training images, and to demonstrate a practical system for gathering sufficient training images.  While most of the techniques we use have been studied in isolation, we believe we are the first to combine them into a complete recognition pipeline that can achieve good results on realistic data.

% We use tools from sparse representation
% to align a test face image with a set of frontal training
% images in the presence of registration error, illumination variation, and partial occlusion. The region
% of attraction of the alignment algorithm is thoroughly characterized on public face datasets such as Multi-PIE.
% We further demonstrate how to capture a set of training images with enough illumination variation that they span test images taken under un-controlled %illumination.
% We have implemented a complete face recognition system, including a projector-based training acquisition system, in order to evaluate how our algorithms %work under practical testing conditions. We show
% that our system can efficiently and effectively recognize faces under
% a variety of realistic conditions, using only
% frontal images under the proposed illuminations as training.

%This paper is enhanced version of a similarly titled conference paper \cite{Wagner2009-CVPR}.  We provide experiment results on a larger internal data set, as well as more extensive experiments on public data sets.  Furthermore, we test several variations on the main pipeline that also achieve good recognition performance.  We also present an alternative formulation of the iterative alignment based on derivatives of the training images rather than the testing images.  We present results with a new internal implementation of L1 minimization using an Augmented Lagrange Method that is significantly faster than our previous interior-point based solver.

We then demonstrate the effectiveness of the proposed new
methods with a complete face recognition system that is {\em
simple, stable, and scalable}. The proposed system performs
robust automatic recognition of subjects from loosely
controlled probe images taken both indoors and outdoors,
using a gallery of
frontal views of the subjects' faces under the proposed
illuminations. An off-the-shelf face
detector\footnote{We use the OpenCV
implementation of the Viola and Jones' face detector
\cite{Viola2004-IJCV}.} is used to detect faces in the test images.

We conduct extensive experiments on the proposed system with
both public databases and a face database that is collected by
our own acquisition system. Our experimental results on
large-scale public face databases show that our algorithm
indeed achieves very good performance on these databases,
exceeding or competing with the state-of-the-art algorithms.
Additionally, our experimental results on our own database
clearly demonstrate that our system not only works well with
images taken under controlled laboratory conditions, but is
capable of handling practical indoor and outdoor illuminations as well.

\noindent{\em Organization of this paper:} In Section \ref{sec:registration},
we derive our robust registration and recognition algorithm within the sparse
representation framework. We elaborate on algorithmic implementation issues,
conduct region of attraction experiments with respect to both 2D in-plane
deformation and 3D pose variation, and discuss its relationship to existing
work. Section \ref{sec:illumination} is dedicated to our training acquisition
system. Using this system, we investigate empirically how many training
illuminations are required to handle practical illumination variations, and
suggest a sufficient set of 38 training illuminations. Extensive experiments on
a large-scale public database and on our own database are conducted in Section
\ref{sec:multipie} and Section \ref{sec:own-data}, respectively, to verify the
proposed system. Section \ref{sec:conclusion} concludes our work with
discussion of promising future directions.

\section{Robust Alignment}\label{sec:registration}
As demonstrated in Figure \ref{fig:promo}(top), the main
limitation of the {\em sparse representation and
classification} (SRC) algorithm of \cite{Wright2009-PAMI} is
the assumption of pixel-accurate alignment between the test
image and the training set. This leads to brittleness under
pose and misalignment, making it inappropriate for deployment
outside a laboratory setting. In this section, we show how this
weakness can be rectified while still preserving the conceptual
simplicity and good recognition performance of SRC.

SRC assumes access to a database of multiple registered
training images per subject, taken under varying illuminations.
The images of subject $i$, stacked as vectors, form a matrix
$A_i \in \Re^{m \times n_i}$. Taken together, all of the images
form a large matrix $A = [ A_1 \mid A_2 \mid \dots \mid A_K ]
\in \Re^{m \times n}$. As argued in \cite{Wright2009-PAMI}, a
well-aligned test image $\y_0$ can be represented as a sparse
linear combination $A \x_0$ of all of the images in the
database,\footnote{We assume the training illuminations are sufficient. We will address how to ensure illumination
sufficiency in the next section.} plus a sparse error $\e_0$
due to corrupted pixels. The sparse representation can be recovered by
minimizing the $\ell^1$-norm\footnote{The $\ell^1$-norm of a
vector, denoted by $\|\cdot\|_1$, is the sum of absolute values of its entries.} of
$\x$ and $\e$:
\begin{equation}
\min_{\x,\e} \, \| \x \|_1 + \| \e\|_1 \quad \subj \quad \y_0 = A \x + \e.
\label{eqn:robust-l1}
\end{equation}
Now suppose that $\y_0$ is subject to some pose or
misalignment, so that instead of observing $\y_0$, we observe
the warped image $\y = \y_0 \circ \tau^{-1}$, for some
transformation $\tau \in T$ where $T$ is a finite-dimensional
group of transformations acting on the image domain.  The
transformed image $\y$ no longer has a sparse representation of
the form $\y = A \x_0 + \e_0$, and naively applying the
algorithm of \cite{Wright2009-PAMI} is no longer appropriate,
as seen in Figure \ref{fig:promo}(top).

\subsection{Batch and Individual Alignment} If the
true deformation $\tau^{-1}$ can be found, then we can apply
its inverse $\tau$ to the test image and it again becomes
possible to find a sparse representation of the resulting
image, as $\y \circ \tau = A \x_0 + \e_0$.\footnote{In the terminology of \cite{baker2004lucas}, this formulation is ``Forward Additive''.}
  This sparsity
provides a strong cue for finding the correct deformation
$\tau$: conceptually, one would like to seek a transformation
$\tau$ that allows the sparsest representation, via
\begin{equation} \label{eqn:L1-L1-conceptual}
\hat{\tau} = \arg\hspace{-2.5mm}\min_{\x,\e,\tau \in T} \| \x \|_1 + \| \e \|_1 \quad \subj \quad \y \circ \tau = A \x + \e.
\end{equation}
For fixed $\tau$, this problem is jointly convex in $\x$ and
$\e$. However, as a simultaneous optimization over the
coefficients $\x$, error representation $\e$, and
transformation $\tau$, it is a difficult, nonconvex
optimization problem. One source of difficulty is the presence
of multiple faces in the matrix $A$:
\eqref{eqn:L1-L1-conceptual} has many local minima that
correspond to aligning $\y$ to different subjects. In this
sense, the misaligned recognition problem differs from the
well-aligned version studied in \cite{Wright2009-PAMI}. For the
well-aligned case, it is possible to directly solve for a
global representation, with no concern for local minima. With
possible misalignment, it is more appropriate to seek the best
alignment of the test face with each subject $i$:
\begin{equation} \label{eqn:per-subject-L1}
\hat \tau_i = \arg\hspace{-2.5mm}\min_{\x,\e,\tau_i \in T} \| \e \|_1 \quad \subj \quad \y \circ \tau_i = A_i \x + \e.
\end{equation}
We no longer penalize $\| \x \|_1$, since $A_i$ consists of
only images of subject $i$ and so $\x$ is no longer expected to
be sparse.

\subsection{Alignment via Sequential $\ell^1$-Minimization} While the problem \eqref{eqn:per-subject-L1} is still nonconvex, for cases of practical interest in face recognition, a good initial guess for the transformation is available, e.g., from the output of a face detector. We can refine this initialization to an estimate of the true transformation by repeatedly linearizing about  the current estimate of $\tau$, and seeking representations of the form:
\begin{equation}
\y\circ \tau + J \Delta \tau = A_i \x + \e.
\end{equation}
Here, $J = \frac{\partial}{\partial \tau} \y \circ \tau$ is the
Jacobian of $\y \circ \tau$ with respect to the transformation
parameters $\tau$, and $\Delta \tau$ is the step in $\tau$. The
above equation is underdetermined if we allow the registration
error $\e$ to be arbitrary. At the correct alignment we
expect the test image to differ from $A_i \x$ only
for the minority of the pixels corrupted by occlusions. Thus,
we seek a deformation step $\Delta \tau$ that best sparsifies the registration error $\e$, in terms of its $\ell^1$-norm:
\begin{equation}
\Delta\hat{\tau}_1 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_1 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e.
\label{eqn:L1-align}
\end{equation}
This is different from the popular choice that
minimizes the $\ell^2$-norm of the registration error:
\begin{equation}
\Delta\hat{\tau}_2 = \arg\hspace{-3.5mm}\min_{\x,\e,\Delta\tau \in T} \| \e \|_2 \quad \subj \quad \y\circ\tau + J \Delta \tau = A_i \x + \e,
\label{eqn:L2-align}
\end{equation}
which is also equivalent to finding the deformation step
$\Delta  \tau$ by solving the least-square problem:
$\min_{\x,\Delta \tau} \|\y \circ \tau + J\Delta \tau - A_i \x
\|_2$. Empirically, we find that if there is only small noise
between $\y_0$ and $A_i\x$, both \eqref{eqn:L1-align} and
$\eqref{eqn:L2-align}$ have similar performance.  However, if
there are occlusions in $\y_0$, sequential
$\ell^1$-minimization \eqref{eqn:L1-align} is significantly
better than sequential $\ell^2$-minimization
\eqref{eqn:L2-align}. Figure \ref{fig:L1-L2-align} shows an
example.

The scheme \eqref{eqn:L1-align} can be viewed as a generalized Gauss-Newton method for minimizing the composition of a nonsmooth objective function (the $\ell^1$-norm) with a differentiable mapping from transformation parameters to transformed images. Such algorithms date at least back to the 1970's  \cite{Cromme1978-NM,Jittorntrum1980-NM}, and continue to attract attention today \cite{Lewis2008-TR}. While space precludes a detailed discussion of their properties, we should mention that the scheme \eqref{eqn:L1-align} is known to converge quadratically in the neighborhood of any local optimum of the $\ell^1$-norm. In practice, this means that $\approx$ 10 to 15 iterations suffice to reach the desired solution. We refer the interested reader to \cite{Jittorntrum1980-NM,Osborne1990-JAMSSB} and the references therein.

\begin{figure}
\centering
{\small
\begin{tabular}{cccc}
\includegraphics[height=.8in]{figures_cvpr/L1_cropped} &
\includegraphics[height=.8in]{figures_cvpr/y_warp_L1} &
\includegraphics[height=.8in]{figures_cvpr/y_hat_L1} &
\includegraphics[height=.8in]{figures_cvpr/e_L1} \\
\includegraphics[height=.8in]{figures_cvpr/L2_cropped} &
\includegraphics[height=.8in]{figures_cvpr/y_warp_L2} &
\includegraphics[height=.8in]{figures_cvpr/y_hat_L2} &
\includegraphics[height=.8in]{figures_cvpr/e_L2} \\
(a) & (b) & (c) & (d)
\end{tabular}
\vspace{0mm}}
\caption{{\bf Comparing alignment of a subject wearing sunglasses by
$\ell^1$ and $\ell^2$ minimization.}
{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
given by the face detector, {\em Red (solid):} Alignment result shown on the same
face; (b) warped testing image using the estimated transformation $\y_0$;
(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
\vspace{-8.0mm}
\end{figure}

In addition to normalizing the training images (which is done
once), it is important to normalize the warped testing image
$\y \circ \tau$ as the algorithm runs.  Without normalization,
the algorithm may fall into a degenerate global minimum
corresponding to zooming in on a dark region of the test
image.  Normalization is done by replacing the linearization of
$\y \circ \tau$ with a linearization of the normalized version
$\tilde \y(\tau) = \frac{\y \circ \tau}{\|\y \circ \tau\|_2}$.
The proposed alignment algorithm can be easily extended to work
in a {\em multiscale} fashion, with benefits both in
convergence behavior and computational cost.  The alignment
algorithm is simply run to completion on progressively less
downsampled versions of the training and testing images, using
the result of one level to initialize the next.

\subsection{Robust Recognition by Sparse Representation} Once
the best transformation $\tau_i$ has been computed for each
subject $i$, the training sets $A_i$ can be aligned to $\y$,
and a global sparse representation problem of the form
\eqref{eqn:robust-l1} can be solved to obtain a discriminative
representation in terms of the entire training set. Moreover,
the per-subject alignment residuals $\| \e \|_1$ can be used to
prune unpromising candidates from the global optimization,
leaving a much smaller and more efficiently solvable problem.
The complete optimization procedure is summarized as Algorithm
\ref{alg:deformable-src}. The parameter $S$ in our algorithm is the number of subjects
considered together to provide a sparse representation for the
test image. If $S = 1$, the algorithm reduces to classification
by registration error; but considering the test image might be
an invalid subject, we typically choose $S = 10$. Since valid
images have a sparse representation in terms of this larger
set, we can reject invalid test images using the {\em sparsity
concentration index} proposed in \cite{Wright2009-PAMI}.
The function $\delta_i(\x)$ in Algorithm \ref{alg:deformable-src}
selects coefficients from the vector $\x$ corresponding to subject $i$.

Another important free parameter in Algorithm
\ref{alg:deformable-src} is the class of deformations $T$. In
our experiments, we typically use 2D similarity
transformations, $T = \mathbb{SE}(2)\times \Re_+$\footnote{Here, SE stands for Special
Euclidean, i.e., 2D rigid transformations.  The $\Re_+$ accounts for the scale.}, for removing alignment error incurred by face
detector, or 2D projective transformations, $T =
\mathbb{GL}(3)\footnote{Here, GL stands for General Linear.
This class of transformations is able to represent distortion
in a perspective image of a planar object.}$, for handling some
pose variation.

In Algorithm \ref{alg:deformable-src}, we also implement a simple heuristic which improves the
performance of our system, based on the observation that the face detector output may be poorly centered on
the face, and may contain a significant amount of the background.  Therefore, before the recognition stage,
instead of aligning the training sets to the
original $\y$ directly obtained from the face detector, we
compute an average transformation $\bar{\tau}$ from $\tau_{k_1},
\tau_{k_2}, \ldots, \tau_{k_S}$ of the top $S$ classes, which
is believed to be better centered, and update $\y$ according to
$\bar{\tau}$. For the 2D similarity transformations, which are
used in our system when initialized by the face detector, a
transformation $\tau$ can be parameterized as $\tau = (\tau^1,
\tau^2, \tau^3, \tau^4)$, where $\tau^1$ and $\tau^2$ represent
the translations in $x$- and $y$-axis, $\tau^3$ represents the
rotation angle and $\tau^4$ represents the scale. Then the
average transformation is simply obtained by taking the
component-wise mean:
\begin{displaymath}
\bar{\tau}^i = (\tau_{k_1}^i + \tau_{k_2}^i + \cdots +
\tau_{k_S}^i) / S, i = 1,2,3,4.
\end{displaymath}
Finally, the training sets are aligned to the new $\y$.

\begin{algorithm}[thb]
\caption{\bf (Deformable Sparse Recovery and Classification for
Face Recognition)} \label{alg:deformable-src}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
$\y\in\Re^m$ and a deformation group $T$.
\STATE
{\bf for} each subject $i$,
\STATE \hspace{3mm} $\tau^{(0)}
\leftarrow I$.
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
\STATE \hspace{6mm}
$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
\STATE
\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$;
\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
residuals $\|\e\|_1$. \STATE Compute an average transformation
$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
$\ell^1$-minimization problem: \hspace{2mm}
$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
\end{small}
\end{algorithmic}
\end{algorithm}
%\vspace{-4mm}

The transformation $\tau$ defines a mapping between the
coordinates of pixels in the large original image and a smaller
(un)warped image. The pixels of the small image are stacked into a vector. To prevent aliasing
artifacts in the downsampled image, one should apply a
smoothing filter to the original image. For a simple implementation, a rectangular window with
regular sampling can used, but in general, the small image need
not be regularly sampled in pixel coordinates.  For example,
the sample locations could be arbitrarily selected from within a
``face shaped'' area. We will discuss how choosing
different windows can affect the performance of our algorithm
in Section \ref{sec:multipie}.

\subsection{System Implementation}
In this section, we discuss the computational issues related to
the implementation of Algorithm~\ref{alg:deformable-src}. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.

@@

\subsection{Experiments on Region of Attraction} We will now present three
experimental results demonstrating the effectiveness of the individual
alignment procedure outlined in the previous section. They show the sufficiency
of the region of attraction, verify effectiveness of the multiscale extension,
and show stability to small pose variations.  We delay large-scale recognition
experiments to Sections \ref{sec:multipie} and \ref{sec:own-data}, after we
have discussed the issue of illumination in the next section.

\noindent {1) \em 2D Deformation.}  We first verify the
    effectiveness of our alignment algorithm with images
    from the CMU Multi-PIE Database \cite{Gross2008-FGR}.
    We select all the subjects in Session 1, use 7
    illuminations per person from Session 1 for training,
    and test on one new illumination from Session
    2.\footnote{The training are illuminations $\{0, 1, 7,
    13, 14, 16, 18\}$ of \cite{Gross2008-FGR}, and the
    testing is the illumination 10. } We manually select
    eye corners in both training and testing as the ground
    truth for registration. We downsample the images to
    $80\times 60$ pixels\footnote{Unless otherwise stated,
    this will be the default resolution at which we prepare
    all our training and testing datasets.} and the distance between the two outer
    eye corners is normalized to be 50 pixels for each
    person. We introduce artificial deformation to the
    testing image with a combination of translation,
    rotation and scaling. We further use the alignment
    error $\|\e\|_1$ as an indicator of success. Let $r_0$
    be the alignment error obtained by aligning a test
    image to the training images without any artificial
    perturbation. When the test image is artificially
    perturbed and aligned, resulting in an alignment error
    $r$, we consider the alignment successful if $|r - r_0 | \leq
    0.01r_0$. Figure \ref{fig:attraction} shows the
    percentage of successful registrations for all subjects
    for each artificial deformation. The results suggest
    that our algorithm works extremely well with
    translation up to 20\% of the eye distance (or 10
    pixels) in all directions and up to $30^\circ$ in-plane
    rotation. We have also tested our alignment algorithm
    with scale variation and it can handle up to 15\%
    change in scale.

We have gathered the statistics of the Viola and Jones'
 face detector on the Multi-PIE dataset. For 4,600 frontal
 images of 230 subjects under 20 different illuminations,
 using manual registration as the ground truth, the average
 misalignment error of the detected faces is about 6 pixels
 and the average variation in scale is 8\%. This falls
 safely inside the region of attraction for our alignment
 algorithm.
\begin{figure*}
\centering
{\small
\begin{tabular}{ccc}
\includegraphics[height=1.2in]{figures_pami/x_y_roa.png} &
\includegraphics[height=1.2in]{figures_pami/y_theta_roa.png} &
\includegraphics[height=1.2in]{figures_pami/y_s_roa.png}\\
(a)&(b)&(c)
\end{tabular}
\begin{tabular}{cccc}
\includegraphics[height=1in]{figures_pami/x_tr.png} &
\includegraphics[height=1in]{figures_pami/y_tr.png} &
\includegraphics[height=1in]{figures_pami/theta.png} &
\includegraphics[height=1in]{figures_pami/scale.png}\\
(d)&(e)&(f)& (g)
\end{tabular}
\vspace{0mm}
}
\caption{{\bf Region of attraction.} Fraction of subjects for which the algorithm successfully aligns a synthetically perturbed test image.  The amount of translation is expressed as a fraction of the distance between the outer eye corners, and the amount of in-plane rotation in degrees. {\bf Top row:} (a) Simultaneous translation in $x$ and $y$ directions. (b) Simultaneous translation in $y$ direction and in-plane rotation. (c) Simultaneous translation in $y$ direction and scale variation. {\bf Bottom row:} (d) Translation in $x$ direction only. (e) Translation in $y$ direction only. (f) In-plane rotation only. (g) Scale variation only.}
\label{fig:attraction}
\vspace{0mm}
\end{figure*}

\noindent{2) \em Multiscale Implementation.}
Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost. Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
We use the same 7 illuminations in
Session 1 as training, and all 20 illuminations in the same
session as testing. We introduce artificial deformation in
both $x$ and $y$ directions up to 16 pixels in the
$80\times 60$ frame, with a step size of 4 pixels, i.e.,
$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
\{-16,-12,\ldots,12,16\}$. We consider an alignment
successful if the estimated coordinates of the eye-corners
are within 1 pixel from the ground truth in the original
image.  In Figure \ref{fig:multiscale}, we report the
alignment success rate, averaged over the artificially
perturbed initial deformations, as a function of the
standard deviation of the Gaussian kernel $\sigma$, for
three choices of the number of scales. As one can see,
using multiscale indeed improves the performance, and when
3 scales are used, a smaller convolution kernel can achieve
a similar performance compared to a much larger kernel when
only 2 scales are used.
\begin{figure}
\centering
\includegraphics[width=3in]{figures_pami/multiscale.png}
\vspace{-2mm}
\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations. A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
\label{fig:multiscale}
\vspace{-4mm}
\end{figure}

\noindent{3) \em 3D Pose Variation.} As densely sampled pose
and
    illumination face images are not available in any of
    the public databases, including Multi-PIE, we have
    collected our own dataset using our own system (to be
    introduced in the next section). We use frontal face
    images of a subject under the 38 illuminations proposed
    in the next section as training. For testing, we
    collect images of the subject under a typical indoor
    lighting condition at pose ranging from $-90^\circ$ to
    $+90^\circ$ with step size 5.625$^\circ$, a total of 33

>>>>

@@

%%%%%%%%% BODY TEXT
\section{L1 Minimization via Augmented Lagrange Multiplier}

\PARstart{I}{n} these supplementary materials we discuss the computational issues related to
the implementation of Algorithm~\ref{alg:deformable-src}, which is
repeated here for convenience. It is
not hard to see that its computational complexity is dominated
by the two steps where the $\ell^1$-norm minimization problems
are solved; namely Step 6 for iterative registration, and Step
14 for global sparse representation.
Fortunately, many fast algorithms for solving these problems have been proposed over the
past ten years. We refer the interested reader to \cite{YangA2010-pp} for a more comprehensive survey of the developments in this area.
That work suggests that \emph{Augmented Lagrange Multiplier}
(ALM) algorithms \cite{Bertsekas1982} strike a good balance
between scalability and accuracy: as first order methods, they
require only lightweight vector operations and matrix-vector
multiplications at each iteration, making them preferable to
more classical solutions such as interior point methods.
However, compared to other first-order methods, they achieve
higher accuracy with a fixed computational budget.
\input{deformable-src.tex}

@@
